Rank verification for exponential families

### 2016-04-14 verifying-winner.tex
Addressed the comments, main result restated

A Schur-concave function is symmetric by default. Conversely a symmetric and log-concave function is Schur-concave, thus many common exponential family distributions lie in this class. Schur-concave function indeed appears a lot, \Cref{sec:applications} includes both continuous and discrete examples.

Now suppose the carrier distribution $g$ in \Cref{eqn:exp_family} is indeed Schur-concave. We hereon call the class of distribution satisfying this {\em Schur-concave exponential family}. Without loss of generality, we can assume $X_1 \ge X_2 \ge \max_{k > 2} X_k$. We will be testing the null hypothesis $H_{01}: \theta_1 \le \max_{j > 1} \theta_j$, which is the union of the null hypotheses $H_{0j}: \theta_1 \le \theta_j$ for $j > 1$.

@@

We prove the ``if'' part first. For any fixed $a$, and $X_1 \ge a$ and $X_2 < a$, we have $X_1 > X_2$ and
$$\exp\left(\theta_1 X_1 + \theta_2 X_2 + \cdots + \theta_n X_n\right) g\left(X\right) h\left(\theta\right) \ge \exp\left(\theta_1 X_2 + \theta_2 X_1 + \cdots + \theta_n X_n\right) g\left(X\right) h\left(\theta\right).$$
Integrating both sides over the region $\left\{X: X_1 \ge a, X_2 < a\right\}$, summing instead if $X$ is a discrete vector, gives
$$\PP\left(X_1 \ge a, X_2 < a\right) \ge \PP\left(X_1 < a, X_2 \ge a\right).$$
Now adding $\PP\left(X_1 \ge a, X_2 \ge a\right)$ to both probabilities gives
$$\PP\left(X_1 \ge a\right) \ge \PP\left(X_2 \ge a\right),$$
meaning that $X_1$ is greater than $X_2$ in stochastic order. The converse follows from swapping the role of $\theta_1$ and $\theta_2$.

@@

\begin{theorem}
Suppose we observe $X = \left(X_1, \ldots, X_n\right)$ from an exponential family with natural parameter $\theta = \left(\theta_1, \ldots, \theta_n\right)$, i.e.
$$X \sim \left(\theta_1 X_1 + \cdots + \theta_n X_n\right) g\left(X\right) h\left(\theta\right),$$
with a Schur-concave carrier measure $g$. Then there is a selective level-$\alpha$ test that verifies $\theta_{\left(1\right)} > \theta_{\left(2\right)}$ by comparing only $X_{\left(1\right)}$ and $X_{\left(2\right)}$. The test also verifies that $X_{\left(1\right)} \ge X_{\left(j\right)}$ stochastically for all $j > 1$. Here $X_{\left(j\right)}$ is the order statistic where $X_{\left(1\right)} \ge \cdots \ge X_{\left(n\right)}$.
\label{thm:main_result}
\end{theorem}

@@

The $p$-value for $H_{0j}$ is thus
\begin{equation}
p_{1j} = \frac{\int_{D_{1j}}^\infty g\left(M_{1j} + z, X_2, \ldots, M_{1j} - z, \ldots, X_n\right) \,dz}{\int_{\max\left\{X_2 - M_{1j}, 0\right\}}^\infty g\left(M_{1j} + z, X_2, \ldots, M_{1j} - z, \ldots, X_n\right) \,dz}.
\label{eqn:p1j}
\end{equation}

@@

For the alternative case where $X_2 \ge M_{13}$, it follows from a similar proof. In either cases, $p_{12} \ge p_{13}$, or in generality, $p_{12} \ge p_{1j}$ for $j \ge 1$. A test based on only $X_1$ and $X_2$ is sufficient for the null hypothesis $H_{0j}$, if the carrying measure $g$ is Schur-concave. By \Cref{thm:stoch}, rejecting this null hypothesis is equivalent to inferring that $X_1 \ge X_j$ for all $j>1$ stochastically.

@@

\begin{theorem}
In the same setup as \Cref{thm:main_result}, there is a selective level-$\alpha$ lower confidence bound for $\theta_{\left(1\right)} - \max_{j>1} \theta_{\left(j\right)}$.
\end{theorem}

@@

Again we start with assuming $X_1 \ge \cdots \ge X_n$. The null hypothesis $H_{01}^\delta$ can be thought of as $\bigcup_{j > 1} H_{0j}^\delta$, where $H_{0j}^\delta$ is $\theta_1 - \theta_j \le \delta$. In other words, we focus on the conditional law
$$\mathcal{L}_{\theta_1 - \theta_j = \delta} \left(D_{1j} \middle| D_{1j}, X_2, \ldots, X_n, A\right),$$
which is the truncated distribution
\begin{align*}
D_{1j} & \sim \exp\left(\left(\theta_1 - \theta_j\right) D_{1j} + \theta_2 X_2 + \cdots + \left(\theta_1 + \theta_j\right) M_{1j} + \cdots + \theta_n X_n \right) \\
& ~~~~~~~~ g\left(M_{ij} + D_{ij}, X_2, \ldots, M_{ij} - D_{ij}, \ldots X_n\right) 1_{A_1} \\
& \sim \exp\left(\delta D_{ij}\right) g\left(M_{ij} + D_{ij}, X_2, \ldots, M_{ij} - D_{ij}, \ldots X_n\right) 1_{A_1}.
\end{align*}

>>>>

A Schur-concave function is symmetric by default since $a \succeq b$ and $b \succeq a$ if and only if $b$ is a permutation of the coordinates of $a$. Conversely a symmetric and log-concave function is Schur-concave \cite{Marshall:2010hb}, thus many common exponential family distributions lie in this class. Schur-concave function indeed appears a lot, \Cref{sec:applications} includes both continuous and discrete examples.

Now suppose the carrier distribution $g$ in \Cref{eqn:exp_family} is indeed Schur-concave. We hereon call the class of distribution satisfying this {\em Schur-concave exponential family}. Without loss of generality, otherwise by relabeling, we consider only the event $X_1 \ge X_2 \ge \max_{k > 2} X_k$. We will be testing the null hypothesis $H_{01}: \theta_1 \le \max_{j > 1} \theta_j$, which is the union of the null hypotheses $H_{01j}: \theta_1 \le \theta_j$ for $j > 1$.

@@

We prove the ``if'' part first. For any fixed $a$, and $X_1 \ge a$ and $X_2 < a$, we have $X_1 > X_2$ and
$$\exp\left(\theta_1 X_1 + \theta_2 X_2 + \cdots + \theta_n X_n\right) g\left(X\right) h\left(\theta\right) \ge \exp\left(\theta_1 X_2 + \theta_2 X_1 + \cdots + \theta_n X_n\right) g\left(X\right) h\left(\theta\right).$$
Integrating both sides over the region $\left\{X: X_1 \ge a, X_2 < a\right\}$ gives
$$\PP\left(X_1 \ge a, X_2 < a\right) \ge \PP\left(X_1 < a, X_2 \ge a\right).$$
Now adding $\PP\left(X_1 \ge a, X_2 \ge a\right)$ to both probabilities gives
$$\PP\left(X_1 \ge a\right) \ge \PP\left(X_2 \ge a\right),$$
meaning that $X_1$ is greater than $X_2$ in stochastic order. The converse follows from swapping the role of $\theta_1$ and $\theta_2$.

@@

\begin{theorem}
Suppose we observe $X = \left(X_1, \ldots, X_n\right)$ from an exponential family with natural parameter $\theta = \left(\theta_1, \ldots, \theta_n\right)$, i.e.
$$X \sim \left(\theta_1 X_1 + \cdots + \theta_n X_n\right) g\left(X\right) h\left(\theta\right),$$
with a Schur-concave carrier measure $g$. Then a naive two-sided test comparing only $X_{j^*}$ and $\max_{j \ne j^*} X_j$, where $j^* = \argmax_{j} X_j$, can verify $\theta_{j^*} \ge \max_{j \ne j^*} \theta_{j}$ with selective level $\alpha$. The test also verifies that $X_{j^*} \ge X_j$ stochastically for all $j \ne j^*$.
\label{thm:main_result}
\end{theorem}

@@

The $p$-value for $H_{01j}$ is thus
\begin{equation}
p_{1j} = \frac{\int_{D_{1j}}^\infty g\left(M_{1j} + z, X_2, \ldots, M_{1j} - z, \ldots, X_n\right) \,dz}{\int_{\max\left\{X_2 - M_{1j}, 0\right\}}^\infty g\left(M_{1j} + z, X_2, \ldots, M_{1j} - z, \ldots, X_n\right) \,dz}.
\label{eqn:p1j}
\end{equation}

@@

For the alternative case where $X_2 \ge M_{13}$, it follows from a similar proof. In either cases, $p_{12} \ge p_{13}$, or in generality, $p_{12} \ge p_{1j}$ for $j > 1$. A test based on only $X_1$ and $X_2$ is sufficient for the null hypothesis $H_{01}$, if the carrying measure $g$ is Schur-concave. Indeed, since the denominator of $p_{12}$ under the null hypothesis, this one-sided test comparing $X_1$ and $X_2$ is identical to a naive two-sided test comparing $X_1$ and $X_2$. By \Cref{thm:stoch}, rejecting this null hypothesis is equivalent to inferring that $X_1 \ge X_j$ for all $j>1$ stochastically.

@@

\begin{theorem}
In the same setup as \Cref{thm:main_result}, there is a selective level-$\alpha$ lower confidence bound for $\theta_{j^*} - \max_{j \ne j^*} \theta_j$.
\end{theorem}

@@

Again we start with assuming $X_1 \ge \cdots \ge X_n$. The null hypothesis $H_{01}^\delta$ can be thought of as $\bigcup_{j > 1} H_{01j}^\delta$, where $H_{01j}^\delta$ is $\theta_1 - \theta_j \le \delta$. In other words, we focus on the conditional law
$$\mathcal{L}_{\theta_1 - \theta_j = \delta} \left(D_{1j} \middle| D_{1j}, X_2, \ldots, X_n, A\right),$$
which is the truncated distribution
\begin{align*}
D_{1j} & \sim \exp\left(\left(\theta_1 - \theta_j\right) D_{1j} + \theta_2 X_2 + \cdots + \left(\theta_1 + \theta_j\right) M_{1j} + \cdots + \theta_n X_n \right) \\
& ~~~~~~~~ g\left(M_{ij} + D_{ij}, X_2, \ldots, M_{ij} - D_{ij}, \ldots X_n\right) 1_{A_1} \\
& \sim \exp\left(\delta D_{ij}\right) g\left(M_{ij} + D_{ij}, X_2, \ldots, M_{ij} - D_{ij}, \ldots X_n\right) 1_{A_1}.
\end{align*}
